---
title: "Investigating the exponential distribution in R"
author: "Kiattisak Chaisomboon"
date: "07/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

We will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with "rexp(n, $\lambda$)" where n is number of observations and $\lambda$ is the rate parameter.

## Simulations

Creating the list of sample mean (named 'mns') and the list of sample variance (named 'vrs') from 1000 samples which simulated using the function of the exponential distribution "rexp(n, $\lambda$)" where n = 40 and $\lambda$ = 0.2
```{r }
library(ggplot2)

set.seed(23)
sim <- 1000
n <- 40
lambda = 0.2

mns = NULL
vrs = NULL
for (i in 1 : sim) {
  rnd = rexp(n, lambda)
  mns = c(mns, mean(rnd))
  vrs = c(vrs, var(rnd))
}
```

## Sample Mean versus Theoretical Mean

Calculating the sample mean.
```{r }
sample.mean <- mean(mns)
sample.mean
```
\
The mean of exponential distribution is $1/\lambda$, can be used to calculate the theoritical mean as the follow:
```{r }
theoritical.mean <- 1/lambda
theoritical.mean
```
\
The sample mean (5.01425) is very close to the theoritical mean (5), as shown in the following figure: 
```{r fig.height=2}
dfm <- data.frame(mns)
ggplot(dfm, aes(mns)) + 
  geom_histogram(bins=30, colour="black", fill="white") +
  geom_vline(aes(xintercept = sample.mean, colour='Sample mean')) + 
  geom_vline(aes(xintercept = theoritical.mean, colour='Theoritical mean')) + 
  scale_color_manual(name="", values = c('Sample mean' = 'red', 
                                    'Theoritical mean' = 'blue')) +
  xlab("Mean") + ylab("Density")
```

## Sample Variance versus Theoretical Variance

Calculating the sample variance.
```{r }
sample.variance <- mean(vrs)
sample.variance
```
\
The variance of exponential distribution is $(1/\lambda)^2$, can be used to calculate the theoritical variance as the follow:
```{r }
theoritical.variance <- (1/lambda)^2
theoritical.variance
```
\
The sample variance (24.92895) is very close to the theoritical variance (25), as shown in the following figure: 
```{r fig.height=2}
dfv <- data.frame(vrs)
ggplot(dfv, aes(vrs)) + 
  geom_histogram(bins=30, colour="black", fill="white") +
  geom_vline(aes(xintercept = sample.variance, colour='Sample variance')) + 
  geom_vline(aes(xintercept = theoritical.variance, colour='Theoritical variance')) + 
  scale_color_manual(name="", values = c('Sample variance' = 'red', 
                                    'Theoritical variance' = 'blue')) +
  xlab("Variance") + ylab("Density")
```

## Distribution

The CLT states that the distribution of averages of iid variables (properly normalized) becomes that of a standard normal as the sample size increases. The result is that $Z = \frac{\overline{X_n} - \mu}{\sigma / \sqrt{n}}$ has a distribution like that of a standard normal for large n.

```{r fig.height=3}
z <- ((mns - sample.mean)/(sample.variance / sqrt(sim)))

df <- data.frame(z)
ggplot(df, aes(x=z)) +
  geom_histogram(aes(y=..density..), bins=30, colour="black", fill="white") +
  geom_density(aes(colour='Sampling Distribution')) + 
  geom_line(stat = "function", fun = "dnorm", args = list(mean = 0, sd = 1),
            aes(colour='Standard Normal Distribution')) +
  scale_color_manual(name="", values = c('Sampling Distribution' = 'red', 
                                    'Standard Normal Distribution' = 'blue')) + 
  xlab("Z") + ylab("Density")
```
