---
title: 'The ToothGrowth data Analysis'
author: "Kiattisak Chaisomboon"
date: "07/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

We are going to analyze the ToothGrowth data in the R datasets package. We will study the change of the tooth growth in pigs depends on supplement type and dose quantity.

## Exploratory Data Analysis

Loading the dataset and check its dimension
```{r }
library(dplyr)
tg <- ToothGrowth
dim(tg)
```
The dataset contains 60 observations and 3 variables are:\
len = Tooth length\
supp = Supplement type (VC or OJ)\
dose = Dose in milligrams/day\
\
Summarizing the basic information from the dataset.
```{r fig.height=3}
summary(tg)
par(mfrow=c(1,2))
boxplot(len ~ supp, data=tg)
boxplot(len ~ dose, data=tg)
```
\
The results show that:\
- Using the orange juice give a tooth growth rate higher than using the acid.\
- Increasing dosage give a higher tooth growth rate.\
We'll set the hypothesis testing related with this information.

## The Change of Tooth Length based on the Type of Supplement

Create variables based on the type of supplement.
```{r }
OJ <- subset(tg, supp == 'OJ')
VC <- subset(tg, supp == 'VC')
```
\
\
We hypothesize that the mean of tooth length when using an orange juice (OC) is higher than the mean of tooth length when using an ascorbic acid (VC), then we set:\
$H_0 : \mu_{oc} = \mu_{vc}$\
$H_{\alpha} : \mu_{oc} > \mu_{vc}$\
Set the confidence interval = 95% ($\alpha = 0.05$)
```{r }
t.test(OJ$len, VC$len, alternative='greater', paired = FALSE, var.equal = FALSE)
```
The output reports that the p-value = 0.03032, is less than $\alpha = 0.05$ then we'll reject the null hypothesis. We are 95% confident that the mean of tooth length when using the orange juice is higher than the mean of tooth length when using the ascorbic acid.

## The Change of Tooth Length based on Level of Dose

Create variables based on level of dose
```{r }
half.dose <- subset(tg, dose == 0.5)
one.dose <- subset(tg, dose == 1)
two.dose <- subset(tg, dose == 2)
```
\
\
We hypothesize that the mean of tooth length when using an one dosage is higher than the mean of tooth length when using a half dosage, then we set:\
$H_0 : \mu_{1.0D} = \mu_{0.5D}$\
$H_{\alpha} : \mu_{1.0D} > \mu_{0.5D}$\
Set the confidence interval = 95% ($\alpha = 0.05$)
```{r }
t.test(one.dose$len, half.dose$len, alternative='greater', paired=FALSE, var.equal=FALSE)
```
The output reports that the p-value = 6.342e-08, is less than $\alpha = 0.05$ then we'll reject the null hypothesis. We are 95% confident that the mean of tooth length when using an one dosage is higher than the mean of tooth length when using a half dosage.\

\
\
Next, we hypothesize that the mean of tooth length when using a double dosage is higher than the mean of tooth length when using an one dosage, then we set:\
$H_0 : \mu_{2.0D} = \mu_{1.0D}$\
$H_{\alpha} : \mu_{2.0D} > \mu_{1.0D}$\
Set the confidence interval = 95% ($\alpha = 0.05$)
```{r }
t.test(two.dose$len, one.dose$len, alternative='greater', paired=FALSE, var.equal=FALSE)
```
The output reports that the p-value = 9.532e-06, is less than $\alpha = 0.05$ then we'll reject the null hypothesis. We are 95% confident that the mean of tooth length when using a double dosage is higher than the mean of tooth length when using an one dosage.\
