---
title: 'The ToothGrowth data Analysis'
author: "Kiattisak Chaisomboon"
date: "07/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

We are going to analyze the ToothGrowth data in the R datasets package. We will study the change of the tooth length in pigs each type of supplement and changing dose quantity.

## Exploratory Data Analysis

First, loading dataset and check its dimension
```{r }
library(dplyr)
tg <- ToothGrowth
dim(tg)
```
The dataset contains 60 rows and 3 columns

Next, summarizing some basic data
```{r fig.height=3}
summary(ToothGrowth)
par(mfrow=c(1,2))
boxplot(len ~ supp, data=tg)
boxplot(len ~ dose, data=tg)
```
It's seem giving orange joice is better than an acid and give more dose is better.

## Inferential Data Analysis

```{r }
OJ <- subset(tg, supp == 'OJ')
VC <- subset(tg, supp == 'VC')
```

OJ greater than VC
```{r }
t.test(OJ$len, VC$len, alternative='greater', paired = FALSE, var.equal = FALSE)
```

```{r }
half.dose <- subset(tg, dose == 0.5)
one.dose <- subset(tg, dose == 1)
two.dose <- subset(tg, dose == 2)
```

One dose effect greater than Half dose
```{r }
t.test(one.dose$len, half.dose$len, alternative='greater', paired=FALSE, var.equal=FALSE)
```

Two dose effect greater than One dose
```{r }
t.test(two.dose$len, one.dose$len, alternative='greater', paired=FALSE, var.equal=FALSE)
```

